Alluvial Diagrams in ggplot2

### 2018-05-17 inst/doc/ggalluvial.rmd
rebuild site

```{r example alluvial diagram using Titanic dataset, echo=FALSE}
ggplot(data = to_lodes(as.data.frame(Titanic),
                       key = "Demographic",
                       axes = 1:3),
       aes(x = Demographic, stratum = stratum, alluvium = alluvium,
           weight = Freq, label = stratum)) +
  geom_alluvium(aes(fill = Survived)) +
  geom_stratum() + geom_text(stat = "stratum") +
  ggtitle("passengers on the maiden voyage of the Titanic",
          "stratified by demographics and survival")
```

@@

```{r alluvia format of Berkeley admissions dataset}
head(as.data.frame(UCBAdmissions), n = 12)
is_alluvial(as.data.frame(UCBAdmissions), logical = FALSE, silent = TRUE)
```

@@

The long format recognized by **ggalluvial** contains _one row per lode_, and can be understood as the result of "gathering" (in the **dplyr** sense) or "pivoting" (in the Microsoft Excel sense) the axis columns of a dataset in the alluvia format into a key-value pair of columns encoding the axis as the key and the stratum as the value. This format requires an additional indexing column that links the rows corresponding to a common cohort, i.e. the lodes of a single alluvium, as illustrated below using the `to_lodes()` defaults on the Berkeley admissions dataset:

```{r lodes format of Berkeley admissions dataset}
UCB_lodes <- to_lodes(as.data.frame(UCBAdmissions), axes = 1:3)
head(UCB_lodes, n = 12)
is_alluvial(UCB_lodes, logical = FALSE, silent = TRUE)
```

@@

```{r time series alluvia diagram of refugees dataset}
data(Refugees, package = "alluvial")
country_regions <- c(
  Afghanistan = "Middle East",
  Burundi = "Central Africa",
  `Congo DRC` = "Central Africa",
  Iraq = "Middle East",
  Myanmar = "Southeast Asia",
  Palestine = "Middle East",
  Somalia = "Horn of Africa",
  Sudan = "Central Africa",
  Syria = "Middle East",
  Vietnam = "Southeast Asia"
)
Refugees$region <- country_regions[Refugees$country]
ggplot(data = Refugees,
       aes(x = year, weight = refugees, alluvium = country)) +
  geom_alluvium(aes(fill = country, colour = country),
                alpha = .75, decreasing = FALSE) +
  scale_x_continuous(breaks = seq(2003, 2013, 2)) +
  theme(axis.text.x = element_text(angle = -30, hjust = 0)) +
  scale_fill_brewer(type = "qual", palette = "Set3") +
  scale_color_brewer(type = "qual", palette = "Set3") +
  facet_wrap(~ region, scales = "fixed") +
  ggtitle("refugee volume by country and region of origin")
```

@@

```{r alluvial diagram of vaccinations dataset}
data(vaccinations)
levels(vaccinations$response) <- rev(levels(vaccinations$response))
ggplot(vaccinations,
       aes(x = survey, stratum = response, alluvium = subject,
           weight = freq,
           fill = response, label = response)) +
  geom_flow() +
  geom_stratum(alpha = .5) +
  geom_text(stat = "stratum", size = 3) +
  theme(legend.position = "none") +
  ggtitle("vaccination survey responses at three points in time")
```

>>>>

```{r example alluvial diagram using Titanic dataset, echo=FALSE}
ggplot(data = to_lodes_form(as.data.frame(Titanic),
                            key = "Demographic",
                            axes = 1:3),
       aes(x = Demographic, stratum = stratum, alluvium = alluvium,
           weight = Freq, label = stratum)) +
  scale_x_discrete(expand = c(.05, .05)) +
  geom_alluvium(aes(fill = Survived)) +
  geom_stratum() + geom_text(stat = "stratum") +
  ggtitle("passengers on the maiden voyage of the Titanic",
          "stratified by demographics and survival")
```

@@

```{r alluvia format of Berkeley admissions dataset}
head(as.data.frame(UCBAdmissions), n = 12)
is_alluvia_form(as.data.frame(UCBAdmissions), axes = 1:3, silent = TRUE)
```

@@

The long format recognized by **ggalluvial** contains _one row per lode_, and can be understood as the result of "gathering" (in the **dplyr** sense) or "pivoting" (in the Microsoft Excel sense) the axis columns of a dataset in the alluvia format into a key-value pair of columns encoding the axis as the key and the stratum as the value. This format requires an additional indexing column that links the rows corresponding to a common cohort, i.e. the lodes of a single alluvium:

```{r lodes format of Berkeley admissions dataset}
UCB_lodes <- to_lodes_form(as.data.frame(UCBAdmissions),
                           axes = 1:3,
                           id = "Cohort")
head(UCB_lodes, n = 12)
is_lodes_form(UCB_lodes, key = x, value = stratum, id = Cohort, silent = TRUE)
```

@@

```{r time series alluvia diagram of refugees dataset}
data(Refugees, package = "alluvial")
country_regions <- c(
  Afghanistan = "Middle East",
  Burundi = "Central Africa",
  `Congo DRC` = "Central Africa",
  Iraq = "Middle East",
  Myanmar = "Southeast Asia",
  Palestine = "Middle East",
  Somalia = "Horn of Africa",
  Sudan = "Central Africa",
  Syria = "Middle East",
  Vietnam = "Southeast Asia"
)
Refugees$region <- country_regions[Refugees$country]
ggplot(data = Refugees,
       aes(x = year, weight = refugees, alluvium = country)) +
  geom_alluvium(aes(fill = country, colour = country),
                alpha = .75, decreasing = FALSE) +
  scale_x_continuous(breaks = seq(2003, 2013, 2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -30, hjust = 0)) +
  scale_fill_brewer(type = "qual", palette = "Set3") +
  scale_color_brewer(type = "qual", palette = "Set3") +
  facet_wrap(~ region, scales = "fixed") +
  ggtitle("refugee volume by country and region of origin")
```

@@

```{r alluvial diagram of vaccinations dataset}
data(vaccinations)
levels(vaccinations$response) <- rev(levels(vaccinations$response))
ggplot(vaccinations,
       aes(x = survey, stratum = response, alluvium = subject,
           weight = freq,
           fill = response, label = response)) +
  scale_x_discrete(expand = c(.05, .05)) +
  geom_flow() +
  geom_stratum(alpha = .5) +
  geom_text(stat = "stratum", size = 3) +
  theme(legend.position = "none") +
  ggtitle("vaccination survey responses at three points in time")
```
